\section{Justification}\label{sec:justification}
This section explains why I have chosen this problematic by describing
my current project, important experiences and my personal interests. \\
I've been working at Sopra Steria for one year, on the project
Chorus Pro, for the Agence pour l'Informatique Financi\`{e}re de l'Etat
(AIFE).
I was in internship in 2018 and I started my apprenticeship in the exact
same team.
It's a big project that handles digital invoices sent by companies to the
State.
It's mostly written in Java, with a Web Portal, REST / Soap API, EDI
exchanges and Spring Batch.
There are a lot of modules, components, environments, teams and of course
a lot of legacy code. \\
Until very recently, there was almost no unit / integration tests, it was
barely manual testing in dev environments and manual functional testing
in staging environment.
The consequences of the lack of testing were obvious and I realized why
testing is so important, the time it can save or the bugs it can prevent.
I saw some people reluctant to add new tests, to automate them or to
improve code quality and it appears that it really is a vicious circle.
\\
\newline
Beside my project where I understood why Quality and Testing are
important, I also attended to, in my opinion the best class of my
training, the Agile class by M. Zam.
This class was a very broad overview of methods, techniques, tools and a
lot of buzzwords that help to create better software.
I would rename this class as Software Craftsmanship, because it is a
bunch of good practices and real life complex methods that are usually
used only by Software Crafters. \\
I was very sceptical by what he showed us, because it seemed very complex
and not natural at all.
Things such as TDD, Pair Programming, DDD were completely different of
what I was used to.
I couldn't understand why people were talking so much about that, how can
people could even work with these methods. \\
So I tried to see what others do, and the best way I found was Meetup.
Meetup are small events, usually with several talks but sometimes you
simply code with other people and you make some tricky exercises or
discover new technologies.
I realized that these unusual methods were actually very common and that
it only depends on you to work with the people you want to. \\
\newline
This problematic was perfectly suited for me. \\
First of all, I'm very interesting in these Quality and Testing concerns,
so I can learn new things, methods, tools about it and see how common
problems are tackled. \\
Second, I was a bit tired of working on the same thing in my project and
I heard that they wanted to work on test automation.
So this problematic could help me to switch to a fresh mission that keeps
me in the mood. \\
Finally, I think that it will be very helpful for me later, in my future
job.
I can learn a lot of new methods and technologies that are used in the
real world and be better prepared to work with.