\section{Relevant Context}\label{sec:relevant-context}
This section describes which contexts are relevant to this problematic
and where the proposed solution is suitable. \\
This problematic is common on non-critical IT projects using
programming languages such as Java, Javascript, Python etc.
By non-critical, we reference any project where there is no human life,
financial market, aerospace system and any other "critical domain" at
the core of the application. \\
There are often a Web UI and an exposed API for the end users.
Each of these interfaces have their dedicated tools for development and
testing.
The management of all the different tools can be quite tricky and
directly impacts the cost of maintenance. \\
These projects can have multiples teams that work on, such as business,
dev or testing.
Some teams usually don't have technical skills, like business or even the
testing team.
The communication between them is often hard and sometimes there is
almost no communication. \\
The methodology used, either Agile or Waterfall, does not matter.
Usually, an Agile project is more likely to follow good practices in
Quality and Testing but this is not necessarily the case.
The proposed solution is not Agile-oriented and will work on Waterfall
projects. \\
Finally, as Quality has a cost, low-quality projects usually have smaller
budget.
Therefore the proposed solution only use community or free licences tools
and try to reuse at most as possible the same elements, tools, data etc.